---
title: "task_2"
author: "LMT"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TASK 2
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(skimr)
```

# 1. Reading Data

```{r}
df_task2 <- read_csv("../data/diabetes.csv")
df_task2
```

## 1.1 Dimension of the Data
```{r}
cat("The dimension of dataset: number of columns", ncol(df_task2), "number of rows: ", nrow((df_task2)))
```

## 1.2 Structure of the dataset

```{r}
str(df_task2) # Structure of dataset
summary(df_task2)
```

```{r}
sapply(df_task2,class)
```
# Before cleaning the dataset

```{r}
# Summary statistics
descriptive_stats <- df_task2 %>%
  summarise(across(
    everything(),
    list(
      mean = ~mean(.x, na.rm = TRUE),
      median = ~median(.x, na.rm = TRUE),
      sd = ~sd(.x, na.rm = TRUE),
      var = ~var(.x, na.rm = TRUE)
    )
  ))

# Display the table
kable(t(descriptive_stats), caption = "Descriptive Statistics (Before Cleaning)")
descriptive_stats

# Plot histogram for each variable
numeric_cols <- names(df_task2)[sapply(df_task2, is.numeric)]

```

```{r}
# Calculate descriptive statistics
desc_stats <- df_task2 %>%
  summarise(across(.cols = everything(), 
                   .fns = list(
                     Mean = ~mean(.x, na.rm = TRUE),
                     Median = ~median(.x, na.rm = TRUE),
                     SD = ~sd(.x, na.rm = TRUE),
                     Variance = ~var(.x, na.rm = TRUE)
                   ),
                   .names = "{.col}_{.fn}")) %>%
  pivot_longer(cols = everything(),
               names_to = c("Variable", "Statistic"),
               names_sep = "_",
               values_to = "Value") %>%
  pivot_wider(names_from = Statistic, values_from = Value)

# Display the table
kable(desc_stats, caption = "Descriptive Statistics of Diabetes Dataset before cleanin", digits = 2)
```

# 2. Cleaning the dataset
## 2.1 Missing value
From the metadata in the Pima info file, some values should not be zero, because it’s biologically impossible.
Those include
Glucose
BloodPressure
SkinThickness	
Insulin	
BMI
```{r}
# Columns where 0 is biologically impossible
zero_na_columns <- c("Glucose", "BloodPressure", "SkinThickness", "Insulin", "BMI")

# Replace 0s with NA in those columns
df_clean <- df_task2
df_clean[zero_na_columns] <- lapply(df_clean[zero_na_columns], function(x) ifelse(x == 0, NA, x))

# Check number of missing values (errors identified)
colSums(is.na(df_clean))

```

```{r}
df_mean_imputed <- df_clean  
# Loop through the columns and fill NA with mean
for (col in zero_na_columns) {
  mean_value <- mean(df_mean_imputed[[col]], na.rm = TRUE)
  df_mean_imputed[[col]][is.na(df_mean_imputed[[col]])] <- mean_value
}

```

## 1.2 duplicate value

```{r}
any(duplicated(df_task2))
```

# After cleaning
```{r}

# Summary statistics
descriptive_stats <- df_mean_imputed %>%
  summarise(across(
    everything(),
    list(mean = ~mean(.x, na.rm = TRUE),
         median = ~median(.x, na.rm = TRUE),
         sd = ~sd(.x, na.rm = TRUE),
         var = ~var(.x, na.rm = TRUE))
  ))

# Display the table
kable(t(descriptive_stats), caption = "Descriptive Statistics (after Cleaning)")

# Plot histogram for each variable
numeric_cols <- names(df_mean_imputed)[sapply(df_mean_imputed, is.numeric)]

for (col in numeric_cols) {
  print(
    ggplot(df_mean_imputed, aes_string(x = col)) +
      geom_histogram(bins = 15, fill = "blue", color = "black", alpha = 0.7) +
      labs(title = paste("Histogram of", col, "before cleaning"),
           x = col,
           y = "Count") +
      theme_minimal()
  )
}
```

# After cleaning compare with the before cleaning
```{r}
# Columns where 0 is biologically impossible
zero_na_columns <- c("Glucose", "BloodPressure", "SkinThickness", "Insulin", "BMI")

# Replace 0s with NA
df_clean <- df_task2
df_clean[zero_na_columns] <- lapply(df_clean[zero_na_columns], function(x) ifelse(x == 0, NA, x))

# Check number of missing values
cat("Missing values (zeros treated as NA for Glucose, BloodPressure, SkinThickness, Insulin, BMI)")
cat(" ")
print(colSums(is.na(df_clean)))

# Impute using mixed methods (mean and median only)
df_imputed <- df_clean
```

```{r}
# Mean imputation for Glucose and BloodPressure (symmetric distributions)
for (col in c("Glucose", "BloodPressure")) {
  mean_value <- mean(df_imputed[[col]], na.rm = TRUE)
  df_imputed[[col]][is.na(df_imputed[[col]])] <- mean_value
}

# Median imputation for SkinThickness, Insulin, and BMI (skewed or potentially correlated distributions)
for (col in c("SkinThickness", "Insulin", "BMI")) {
  median_value <- median(df_imputed[[col]], na.rm = TRUE)
  df_imputed[[col]][is.na(df_imputed[[col]])] <- median_value
}
```


```{r}
# Check for zero values in Insulin after imputation
zero_count_insulin <- sum(df_imputed$Insulin == 0, na.rm = TRUE)
cat("Number of zero values in Insulin after imputation:", zero_count_insulin, "\n")
```


```{r}
# Descriptive statistics after cleaning
descriptive_stats_clean <- df_imputed %>%
  summarise(across(
    everything(),
    list(mean = ~mean(.x, na.rm = TRUE),
         median = ~median(.x, na.rm = TRUE),
         sd = ~sd(.x, na.rm = TRUE),
         var = ~var(.x, na.rm = TRUE))
  ))

kable(descriptive_stats_clean, caption = "Descriptive Statistics After Cleaning")

# Plot histograms for imputed data
numeric_cols <- names(df_imputed)[sapply(df_imputed, is.numeric)]
for (col in numeric_cols) {
  print(
    ggplot(df_imputed, aes_string(x = col)) +
      geom_histogram(bins = 15, fill = "blue", color = "black", alpha = 0.7) +
      labs(title = paste("Histogram of", col, "After Cleaning"),
           x = col,
           y = "Count") +
      theme_minimal()
  )
}
```

```{r}
p_after <- ggplot(df_imputed, aes_string(x = col)) +
  geom_histogram(aes(y = ..count..), bins = 15, fill = "blue", color = "black", alpha = 0.7) +
  stat_bin(bins = 15, geom = "text", aes(label = ..count..), vjust = -0.5, size = 3.5) +
  labs(title = paste("After Data", "Cleaning"), x = col, y = "Count") +
  theme_minimal()
```

```{r}
sum(df_imputed$Insulin < 20)
```
